Double White Dwarf Mergers on Adaptive Meshes\\ I. Methodology and Code Verification

### 2015-08-18 papers/wdmerger_I/paper.tex
Rewording of the section on why we cannot integrate the work done to get the change in gravitational energy.

It is worth pointing out that while \autoref{eq:grav_energy_conservation_update} was correctly obtained 
by \cite{arepo} (whose work was the primary inspiration for this approach), there is a subtle error in that derivation. 
Their argument was that if we integrate the work done over the volume of the zone, this volume integral can be 
converted into a surface integral by use of the divergence theorem (see their Equation 93). Then we can simply 
evaluate the work done at each zone face. However, this application of the divergence theorem appears to be 
mathematically incorrect, despite the fact that this is very similar in spirit to the approach discussed here. 
The author fortuitously obtains the correct factor of $1/2$ corresponding to the change in gravitational self-energy.
However, some reflection reveals that there should be no way to derive this double-counting factor simply by integrating 
the work done: that factor is derived from the unique nature of the gravitational potential, which is generated 
by the same distribution that it exerts a force on for a simulation with self-gravity, and it has to be added in 
intentionally to ensure conservation of total energy for these particular types of potentials. We can apply 
the approach of considering energy transfer between the gas energy and the potential energy to any potential field, 
whether or not it is static, changing in time, or generated by the fluid distribution itself. 
Indeed, we will do so for the rotational potential in \autoref{sec:rotation}, where the 
factor of $1/2$ does not appear. However, it is important also to note that this is not related to a 
global flow of energy or the global nature of the gravitational potential, as claimed by \cite{marcello:2012}. 
This double-counting factor would apply even if there were only two point masses on the grid, since 
work would only be required to bring in the second mass from infinity; the first one can be 
placed at no energy cost. This transfer of energy is explicitly \textit{local} because the 
mass transfer is local. It is true that the mass transfer changes the potential globally, 
but that is implicitly accounted for when using the time-centered potential, and no 
additional steps need to be taken.

>>>>

It is worth pointing out that while \autoref{eq:grav_energy_conservation_update} was correctly obtained 
by \cite{arepo} (whose work was the primary inspiration for this approach), there is a subtle error in that derivation. 
Their argument was that if we integrate the work done over the volume of the zone, this volume integral can be 
converted into a surface integral by use of the divergence theorem (see their Equation 93). Then we can simply 
evaluate the work done at each zone face. However, this application of the divergence theorem is 
mathematically incorrect, despite the fact that this is similar in spirit to the approach discussed here. 
Moreover, using the divergence theorem to integrate the work done in a zone can never obtain 
the factor of $1/2$ corresponding to the change in gravitational self-energy. That factor can be thought of
as a double-counting factor that ensures that we correctly capture the energy of a pairwise interaction. 
When one integrates the work done in the zone, there is a contribution to the energy that cannot be 
expressed as a full time derivative or the divergence of a flux. This reflects the 
global nature of the gravitational potential and means that we cannot get the correct energy expression 
unless we integrate over the entire mass distribution; see \citet[Chapter 4]{shu:1992} and \citet{marcello:2012}. 
However, this approach is still valid when considering energy transfer between the gas energy and a  
potential field, if the change in energy is explicitly local. In the case of gravity, the mass flux
that generates a change in energy is local, but the change in energy is stored in the entire gravitational 
field, and is only captured by a global integral because of the long-range nature of the potential.
